---
title: "Project"
author: "Sateesh"
date: "3/4/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loadpackages, message=FALSE}
if(!require("pacman")) install.packages("pacman")
pacman::p_load(e1071, ggplot2, caret, rmarkdown, corrplot , tidyverse , readxl , dplyr , plyr , randomForest,RcppEigen,ranger,nnet,onehot)
```
Importing the dataset:

```{r}
matches <- read.csv("T20Fixtures.csv")
```

Data Manipulation:

```{r}
colnames(matches)[1] <- c("Date")
matches <- matches[,-c(1)]
```

Removing the records with no result 

```{r}
str(matches)
matches$Countries <- as.character(matches$Countries)
matches$Result <- as.character(matches$Result)
col <- matches[matches$Result == "Match abandoned" | matches$Result == "No result" | matches$Result == "Match Tied" ,]
matches <- anti_join(matches,col)
```

Performing string operations to extract the playing two countries from a single column and also the winner 

```{r}
out <- strsplit(matches$Countries , "v. " , fixed =TRUE)
out <- unlist(out)
matches$team1<- out[c(seq(1,length(out),2))]
matches$team2 <- out[c(seq(2,length(out),2))]
out1 <- strsplit(matches$Result , "won" )
out1 <- unlist(out1)
matches$team1 <- str_trim(matches$team1)
matches$team2 <- str_trim(matches$team2)
matches$winner <- out1[c(seq(1,length(out),2))]
matches <- matches[,c(2,4,5,6)]
matches$winner <- str_trim(matches$winner)
```

EDA

Checking the number of matches that are played by each team:

```{r}
matches$team1 <- as.factor(matches$team1)
matches$team2 <- as.factor(matches$team2)
matches$winner <- as.factor(matches$winner)
```

Number of matches each team has won: 

```{r}
ggplot(data = matches , aes( x = winner))+
  geom_bar()+
  theme(axis.text.x=element_text(size=8, angle=90))+
  xlab("Team")+
  ylab("Number of matches")+
  ggtitle("Matches Won by each team")
```
Winning percentage of each team:

```{r}
team1_matches <- as.data.frame(table(matches$team1))
team2_matches <- as.data.frame(table(matches$team2))
colnames(team1_matches) <- c("team","match1")
colnames(team2_matches) <- c("team","match2")
#team1_matches <- merge(x = team1_matches,y = team2_matches, by = "team" ,all = TRUE)
 total_matches <- merge(team1_matches,team2_matches , by = "team" , all = TRUE)
 total_matches[is.na(total_matches)] <- 0
 total_matches$match1 <- total_matches$match1+total_matches$match2
 total_matches <- total_matches[,c(1,2)]
 colnames(total_matches) <- c("Team" , "MatchesPlayed")
 total_matches_won <- as.data.frame(table(matches$winner))
 colnames(total_matches_won) <- c("Team","MatchesWon")
 Final_Stats <- merge(total_matches , total_matches_won , by = "Team" , all = TRUE)
 Final_Stats[is.na(Final_Stats)] <- 0
```

```{r}
Final_Stats$Per <- round((Final_Stats$MatchesWon / Final_Stats$MatchesPlayed) * 100) 
ggplot(data = Final_Stats , aes( x = Team , y = Per))+
  geom_bar(stat = "identity")+
  theme(axis.text.x=element_text(size=9, angle=90))+
  xlab("Team")+
  ylab("Percentage")+
  ggtitle("Teams with their winning percentage")
```
Teams that played more than 30 matches:

```{r}
Final_Stats$Per <- round((Final_Stats$MatchesWon / Final_Stats$MatchesPlayed) * 100) 
ggplot(data = Final_Stats[Final_Stats$MatchesPlayed >= 30,] , aes( x = Team , y = Per))+
  geom_bar(stat = "identity")+
  theme(axis.text.x=element_text(size=9, angle=90))+
  xlab("Team")+
  ylab("Percentage")+
  ggtitle("Teams with their winning percentage")

```
MOst matches played between two teams:

```{r}
team_matches <- as.data.frame(table(matches$team1,matches$team2))
team_matches <- team_matches[team_matches$Freq != 0,]
colnames(team_matches) <- c("Team1", "Team2" , "Matches Played") 
team_matches[order(-team_matches$`Matches Played`),]
```
Checking the number of matches won by teams in the grounds
```{r}
ground_stats <- data.frame(table(matches$Ground,matches$winner))
ggplot(data = ground_stats[ground_stats$Var2 == "India" & ground_stats$Freq >1  ,],aes(x = Var1 , y = Freq))+
  geom_bar(stat = "identity")+
  theme(axis.text.x=element_text(size=8, angle=90))+
  ylab("No of matches played")+
  xlab("Ground")+
  ggtitle("Matches played by India at respective grounds")
```
Loading the File having the details regarding the countries ground belong to  
```{r}
ground_countries <- read.csv("Ground_details.csv")
head(ground_countries)
names(ground_countries)[1] <- "Ground"
head(ground_countries)
```


```{r}
matches1 <- matches %>% left_join(ground_countries,by = ("Ground"))
str(matches1)
```


```{r}
matches1$team1 <- as.character(matches1$team1)
matches1$team2 <- as.character(matches1$team2)
matches1$Country <- as.character(matches1$Country)
```

Creating a new variable called home_ground which would be 0 for away game and 1 for home game(team1 is considered as base team here)
```{r}
matches1$home_ground <- 0
matches1$home_ground[matches1$team1 == matches1$Country] <- 1
matches1$home_ground[matches1$team2 == matches1$Country] <- 2
```
```{r}
matches1 <- matches1[,-c(1,5)]
```
```{r}
##home_away_stats <-(table(matches1$winner,matches1$home_ground))
```

Below is the final dataframe which would be used for Machine learning
```{r}
head(matches1)
str(matches1)
matches1$team1 <- as.factor(matches1$team1)
matches1$team2 <- as.factor(matches1$team2)
matches1$home_ground <- as.factor(matches1$home_ground)
```
```{r}
matches2 <- matches1
```

Feature engineering :
```{r}
sum(is.na(matches1))
dmy <- dummyVars(~team1, data = matches1)
trsf1 <- data.frame(predict(dmy, newdata = matches1))
dmy <- dummyVars(~team2, data = matches1)
trsf2 <- data.frame(predict(dmy, newdata = matches1))
matches1 <- cbind(matches1,trsf1,trsf2)
matches1 <- matches1[,c(-1,-2)]
```
```{r}
set.seed(123)
trainindex <- createDataPartition(matches1$winner, p=0.8, list= FALSE)
matches_train <- matches1[trainindex, ]
matches_test <- matches1[-trainindex, ]
```
### SVM Linear deafult
```{r}
svm_linear <- svm(winner~., data = matches_train , cost = 0.01 , kernel = "linear")
summary(svm_linear)
```
```{r}
pred <- predict(svm_linear,matches_train)
conf.matrix1 <- table(pred,matches_train$winner)
accuracy_svm_linear <- (sum(diag(conf.matrix1))) / sum(conf.matrix1)
accuracy_svm_linear
```
### Best Linear SVM
```{r}
tunesvm2 <- tune(svm,winner~., scale = FALSE,data = matches_train, kernel = "linear" ,
     ranges = list(epsilon = seq(0,1,0.25), cost = 2^(2:5) , degree = seq(2,5)))
```
```{r}
svm_linear <- tunesvm2$best.model
summary(svm_linear)
```

```{r}
bestpred1 <- predict(svm_linear, matches_train)
conf.matrix1 <- table(bestpred1,matches_train$winner)
svm_linear_train_acc <- (sum(diag(conf.matrix1))) / sum(conf.matrix1)
test_pred <- predict(svm_linear, matches_test)
conf.matrix2 <- table(test_pred,matches_test$winner)
svm_linear_test_acc <- (sum(diag(conf.matrix2))) / sum(conf.matrix2)
round(svm_linear_train_acc,3)
round(svm_linear_test_acc,3)
```
### Polynomial SVM
```{r}
tunesvm2 <- tune(svm,winner~., scale = FALSE,data = matches_train, kernel = "polynomial" ,
     ranges = list(epsilon = seq(0,1,0.25), cost = 2^(2:5) , degree = seq(2,5)))
```
```{r}
tunesvm2$best.parameters
bestsvm2 <- tunesvm2$best.model
summary(bestsvm2)
bestpred2 <- predict(bestsvm2, matches_train)
```
```{r}
conf.matrix3 <- table(Predicted = bestpred2, Actual = matches_train$winner)
accuracy3 <- (sum(diag(conf.matrix3))) / sum(conf.matrix3)
accuracy3
```
### Radial SVM
```{r}
tunesvm3 <- tune(svm,winner~., scale = FALSE,data = matches_train, kernel = "radial" ,
     ranges = list(epsilon = seq(0,1,0.25), cost = 2^(2:7)))
```
```{r}
radial_svm <- tunesvm3$best.model
summary(radial_svm)
```
Radial kernel accuracy
```{r}
bestpred3 <- predict(radial_svm, matches_train)
conf.matrix4 <- table(Predicted = bestpred3, Actual = matches_train$winner)
radial_svm_train_acc <- (sum(diag(conf.matrix4))) / sum(conf.matrix4)
radial_test_pred <- predict(radial_svm,matches_test)
conf.matrix5 <- table(Predicted = radial_test_pred, Actual = matches_test$winner)
radial_svm_test_acc <- (sum(diag(conf.matrix5))) / sum(conf.matrix5)
radial_svm_train_acc
radial_svm_test_acc
```
### Random Forests
```{r}
rf_fit <- train(winner ~ ., 
                data = matches_train, 
                method = "ranger")

```
Random Forests Train
```{r}
train_pred <- predict(rf_fit,matches_train)
conf.matrix4 <- table(Predicted = train_pred, Actual = matches_train$winner)
rf_train_acc <- (sum(diag(conf.matrix4))) / sum(conf.matrix4)
test_pred <- predict(rf_fit,matches_test)
conf.matrix4 <- table(Predicted = test_pred, Actual = matches_test$winner)
rf_test_acc <- (sum(diag(conf.matrix4))) / sum(conf.matrix4)
rf_train_acc
rf_test_acc
```
### Evaluating the results
```{r}
models <- data.frame("models" = c("SVM Linear","SVM Radial","Random Forests") , "Train accuracy" = c(svm_linear_train_acc,radial_svm_train_acc,rf_train_acc) , "Test accuracy" = c(svm_linear_test_acc,radial_svm_test_acc,rf_test_acc))
models
```
Testing on the world cup fixtures
Stage 1 

Importing the second worsksheet of 2020 t20 worldcup fixtures
```{r}
fixtures <- read_excel("World_cup_Fixtures.xlsx")
```
```{r}
fixtures$Team1 <- str_trim(fixtures$Team1)
fixtures$Team2 <- str_trim(fixtures$Team2)
teams <- distinct(fixtures,fixtures$Team1)
teams1 <- distinct(fixtures,fixtures$Team2)
names(teams1) <- names(teams)
teams <- rbind(teams , teams1)
names(teams) <- "team"
teams <- distinct(teams)
teams <- teams[teams$team != "A1" & teams$team !="A2" & teams$team !="B1" & teams$team !="B2",]
par_teams <- teams$team
```
Predicting winners from Qualifiers
```{r}
QF <- fixtures %>% filter(Round == "Qualifier A" | Round == "Qualifier B")
#QFA <- QFA[,c(2,3)]
str(QF)
QF$home_ground <- 0
QF$home_ground[QF$Team1 == "Australia"] <- 1
QF$home_ground[QF$Team2 == "Australia"] <- 1
head(QF)
QF$Team1 <- as.factor(QF$Team1)
QF$Team2 <- as.factor(QF$Team2)
QF$home_ground <- as.factor(QF$home_ground)
head(QF)
colnames(QF) <- c("Date","team1","team2","Round","home_ground")
QF$team1 <- str_trim(QF$team1)
QF$team2 <- str_trim(QF$team2)
```



```{r}
QFA1 <- QF[,c(2,3,5)]
QFA1$home_ground <- factor(QFA1$home_ground,levels = c(0,1))
QFA1$team1 <- factor(QFA1$team1, levels=c("Afghanistan","Argentina","Australia","Austria","Bahrain","Bangladesh","Belgium","Belize","Bermuda","Bhutan","Botswana","Brazil","Bulgaria","Canada","Cayman Islands","Chile","Costa Rica","Czech Republic","Denmark","England","Finland","Germany","Ghana","Gibraltar","Greece","Guernsey","Hong Kong","Hungary","ICC World XI","India","Ireland","Italy","Jersey","Kenya","Kuwait","Luxembourg","Malawi","Malaysia","Maldives","Malta","Mexico","Namibia","Nepal","Netherlands","New Zealand","Nigeria","Oman","Pakistan","Papua New Guinea","Peru","Philippines","Qatar","Romania","Samoa","Scotland","Singapore","South Africa","Spain","Sri Lanka","Uganda","United Arab Emirates","West Indies","Zimbabwe"
))
QFA1$team2 <- factor(QFA1$team2, levels=c("Afghanistan","Argentina","Australia","Austria","Bangladesh","Belize","Bermuda","Bhutan","Botswana","Brazil","Bulgaria","Canada","Cayman Islands","Chile","Costa Rica","Czech Republic","Denmark","England","Finland","Germany","Ghana","Gibraltar","Hong Kong","Hungary","ICC World XI","Iceland","India","Ireland","Italy","Jersey","Kenya","Kuwait","Luxembourg","Malaysia","Maldives","Malta","Mexico","Mozambique","Namibia","Nepal","Netherlands","New Zealand","Nigeria","Norway","Oman","Pakistan","Panama","Papua New Guinea","Philippines","Portugal","Qatar","Saudi Arabia","Scotland","Serbia","Singapore","South Africa","Spain","Sri Lanka","Thailand","Turkey","United Arab Emirates","United States","Vanuatu","West Indies","Zimbabwe"))
```

```{r}
dmy <- dummyVars(~team1, data = QFA1)
trsf1 <- data.frame(predict(dmy, newdata = QFA1))
dmy <- dummyVars(~team2, data = QFA1)
trsf2 <- data.frame(predict(dmy, newdata = QFA1))
QFA1 <- cbind(QFA1,trsf1,trsf2)
QFA1 <- QFA1[,c(-1,-2)]
```

```{r}
predictions <- predict(rf_fit,QFA1)
QF$predictions <- predictions
```
```{r}
QF1 <- QF %>% filter(QF$Round == "Qualifier A")
QF1 <- data.frame(table(QF1$predictions))
QF1 <- QF1 %>% arrange(desc(Freq))
A1 <- QF1$Var1[1]
A2 <- QF1$Var1[2]
QF2 <- QF %>% filter(QF$Round == "Qualifier B")
QF2 <- data.frame(table(QF2$predictions))
QF2 <- QF2 %>% arrange(desc(Freq))
B1 <- QF2$Var1[1]
B2 <- QF2$Var1[2]
A1 <- as.character(A1)
A2 <- as.character(A2)
B2 <- as.character(B2)
B1 <- as.character(B1)

```
Srilanka and Ireland advance to Super12 from Qualifier A
Bangladesh and Namibia advance to Super12 from Qualifier B

##### Advancing to Super 12 
```{r}
QF <- fixtures %>% filter(Round == "Super12-1" | Round == "Super12-2")
QF[QF$Team1 == "A1",]$Team1 <- A1
QF[QF$Team1 == "A2",]$Team1 <- A2
QF[QF$Team2 == "A1",]$Team2 <- A1
QF[QF$Team2 == "A2",]$Team2 <- A2
QF[QF$Team2 == "B1",]$Team2 <- B1
QF[QF$Team2 == "B2",]$Team2 <- B2
QFA <- QF[,c(2,3)]
str(QFA)
QFA$home_ground <- 0
QFA$home_ground[QFA$Team1 == "Australia"] <- 1
QFA$home_ground[QFA$Team2 == "Australia"] <- 1
head(QFA)
QFA$Team1 <- as.factor(QFA$Team1)
QFA$Team2 <- as.factor(QFA$Team2)
QFA$home_ground <- as.factor(QFA$home_ground)
head(QFA)
colnames(QFA) <- c("team1","team2","home_ground")
QFA$team1 <- str_trim(QFA$team1)
QFA$team2 <- str_trim(QFA$team2)
```

```{r}
QFA$home_ground <- factor(QFA$home_ground,levels = c(0,1))
QFA$team1 <- factor(QFA$team1, levels=c("Afghanistan","Argentina","Australia","Austria","Bahrain","Bangladesh","Belgium","Belize","Bermuda","Bhutan","Botswana","Brazil","Bulgaria","Canada","Cayman Islands","Chile","Costa Rica","Czech Republic","Denmark","England","Finland","Germany","Ghana","Gibraltar","Greece","Guernsey","Hong Kong","Hungary","ICC World XI","India","Ireland","Italy","Jersey","Kenya","Kuwait","Luxembourg","Malawi","Malaysia","Maldives","Malta","Mexico","Namibia","Nepal","Netherlands","New Zealand","Nigeria","Oman","Pakistan","Papua New Guinea","Peru","Philippines","Qatar","Romania","Samoa","Scotland","Singapore","South Africa","Spain","Sri Lanka","Uganda","United Arab Emirates","West Indies","Zimbabwe"
))
QFA$team2 <- factor(QFA$team2, levels=c("Afghanistan","Argentina","Australia","Austria","Bangladesh","Belize","Bermuda","Bhutan","Botswana","Brazil","Bulgaria","Canada","Cayman Islands","Chile","Costa Rica","Czech Republic","Denmark","England","Finland","Germany","Ghana","Gibraltar","Hong Kong","Hungary","ICC World XI","Iceland","India","Ireland","Italy","Jersey","Kenya","Kuwait","Luxembourg","Malaysia","Maldives","Malta","Mexico","Mozambique","Namibia","Nepal","Netherlands","New Zealand","Nigeria","Norway","Oman","Pakistan","Panama","Papua New Guinea","Philippines","Portugal","Qatar","Saudi Arabia","Scotland","Serbia","Singapore","South Africa","Spain","Sri Lanka","Thailand","Turkey","United Arab Emirates","United States","Vanuatu","West Indies","Zimbabwe"))
```

```{r}
dmy <- dummyVars(~team1, data = QFA)
trsf1 <- data.frame(predict(dmy, newdata = QFA))
dmy <- dummyVars(~team2, data = QFA)
trsf2 <- data.frame(predict(dmy, newdata = QFA))
QFA <- cbind(QFA,trsf1,trsf2)
QFA <- QFA[,c(-1,-2)]
```

```{r}
predictions <- predict(rf_fit,QFA)
QF$predictions <- predictions
```

```{r}
QF1 <- QF %>% filter(QF$Round == "Super12-1")
QF1 <- data.frame(table(QF1$predictions))
QF1 <- QF1 %>% arrange(desc(Freq))
QF2 <- QF %>% filter(QF$Round == "Super12-2")
QF2 <- data.frame(table(QF2$predictions))
QF2 <- QF2 %>% arrange(desc(Freq))
S1 <- as.character(QF1$Var1[1])
S2 <- as.character(QF1$Var1[2])
S3 <- as.character(QF2$Var1[1])
S4 <- as.character(QF2$Var1[2])
```
After Super 12 From Group 1 and Group 2 :
Australia and Pakistan Advance to Semi finals
India and England Advance to Semi FInals

Since the stats are not yet out we assume that the first Semifinalist from group 1 play with the second Semifinalist from group 2 
SO now 
First Semifinal would be between 
Australia and England
Second Semifinal would be between
India and Pakistan

### Semi Finals

```{r}
Semis <- data.frame("team1" = c("Australia","Pakistan"),"team2" = c("England","India"), "home_ground" = c(1,0))
```

```{r}
QFA <- Semis
QFA$home_ground <- factor(QFA$home_ground,levels = c(0,1))
QFA$team1 <- factor(QFA$team1, levels=c("Afghanistan","Argentina","Australia","Austria","Bahrain","Bangladesh","Belgium","Belize","Bermuda","Bhutan","Botswana","Brazil","Bulgaria","Canada","Cayman Islands","Chile","Costa Rica","Czech Republic","Denmark","England","Finland","Germany","Ghana","Gibraltar","Greece","Guernsey","Hong Kong","Hungary","ICC World XI","India","Ireland","Italy","Jersey","Kenya","Kuwait","Luxembourg","Malawi","Malaysia","Maldives","Malta","Mexico","Namibia","Nepal","Netherlands","New Zealand","Nigeria","Oman","Pakistan","Papua New Guinea","Peru","Philippines","Qatar","Romania","Samoa","Scotland","Singapore","South Africa","Spain","Sri Lanka","Uganda","United Arab Emirates","West Indies","Zimbabwe"
))
QFA$team2 <- factor(QFA$team2, levels=c("Afghanistan","Argentina","Australia","Austria","Bangladesh","Belize","Bermuda","Bhutan","Botswana","Brazil","Bulgaria","Canada","Cayman Islands","Chile","Costa Rica","Czech Republic","Denmark","England","Finland","Germany","Ghana","Gibraltar","Hong Kong","Hungary","ICC World XI","Iceland","India","Ireland","Italy","Jersey","Kenya","Kuwait","Luxembourg","Malaysia","Maldives","Malta","Mexico","Mozambique","Namibia","Nepal","Netherlands","New Zealand","Nigeria","Norway","Oman","Pakistan","Panama","Papua New Guinea","Philippines","Portugal","Qatar","Saudi Arabia","Scotland","Serbia","Singapore","South Africa","Spain","Sri Lanka","Thailand","Turkey","United Arab Emirates","United States","Vanuatu","West Indies","Zimbabwe"))
```

```{r}
dmy <- dummyVars(~team1, data = QFA)
trsf1 <- data.frame(predict(dmy, newdata = QFA))
dmy <- dummyVars(~team2, data = QFA)
trsf2 <- data.frame(predict(dmy, newdata = QFA))
QFA <- cbind(QFA,trsf1,trsf2)
QFA <- QFA[,c(-1,-2)]
```

```{r}
predictions <- predict(rf_fit,QFA)
Semis$predictions <- predictions
```

According to our predictions Australia and India advance to Finals 

```{r}
final_match <- data.frame("team1" = c("Australia"),"team2" = c("India"), "home_ground" = c(1))
```

```{r}
QFA <- final_match
QFA$home_ground <- factor(QFA$home_ground,levels = c(0,1))
QFA$team1 <- factor(QFA$team1, levels=c("Afghanistan","Argentina","Australia","Austria","Bahrain","Bangladesh","Belgium","Belize","Bermuda","Bhutan","Botswana","Brazil","Bulgaria","Canada","Cayman Islands","Chile","Costa Rica","Czech Republic","Denmark","England","Finland","Germany","Ghana","Gibraltar","Greece","Guernsey","Hong Kong","Hungary","ICC World XI","India","Ireland","Italy","Jersey","Kenya","Kuwait","Luxembourg","Malawi","Malaysia","Maldives","Malta","Mexico","Namibia","Nepal","Netherlands","New Zealand","Nigeria","Oman","Pakistan","Papua New Guinea","Peru","Philippines","Qatar","Romania","Samoa","Scotland","Singapore","South Africa","Spain","Sri Lanka","Uganda","United Arab Emirates","West Indies","Zimbabwe"
))
QFA$team2 <- factor(QFA$team2, levels=c("Afghanistan","Argentina","Australia","Austria","Bangladesh","Belize","Bermuda","Bhutan","Botswana","Brazil","Bulgaria","Canada","Cayman Islands","Chile","Costa Rica","Czech Republic","Denmark","England","Finland","Germany","Ghana","Gibraltar","Hong Kong","Hungary","ICC World XI","Iceland","India","Ireland","Italy","Jersey","Kenya","Kuwait","Luxembourg","Malaysia","Maldives","Malta","Mexico","Mozambique","Namibia","Nepal","Netherlands","New Zealand","Nigeria","Norway","Oman","Pakistan","Panama","Papua New Guinea","Philippines","Portugal","Qatar","Saudi Arabia","Scotland","Serbia","Singapore","South Africa","Spain","Sri Lanka","Thailand","Turkey","United Arab Emirates","United States","Vanuatu","West Indies","Zimbabwe"))
```

```{r}
dmy <- dummyVars(~team1, data = QFA)
trsf1 <- data.frame(predict(dmy, newdata = QFA))
dmy <- dummyVars(~team2, data = QFA)
trsf2 <- data.frame(predict(dmy, newdata = QFA))
QFA <- cbind(QFA,trsf1,trsf2)
QFA <- QFA[,c(-1,-2)]
```

```{r}
predictions <- predict(rf_fit,QFA)
final_match$predictions <- predictions
predictions
```
India is predicted as the winner 







